### 2021-09-27 _targets.Rmd
add toy example sketch

---
title: "Supplementary information"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Setup

Set up the workflow pipeline and options. We first load the `targets` package and remove the potentially outdated workflow.

```{r}
library(targets)
library(tibble)
library(tidyr)
library(dplyr)
library(scoringutils)
library(ggplot2)
library(purrr)
library(here)
tar_unscript()
```

We now define shared global options across our workflow and load R functions from the `R` folder.

```{targets globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
library(purrr)
library(here)
functions <- list.files(here("R"), full.names = TRUE)
walk(functions, source)
tar_option_set(
  packages = c("tibble", "tidyr", "dplyr", "scoringutils", "ggplot2", "purrr")
)
```

# Toy example

As a toy example we simulate a grid of data with varying exponential growth rates and a Poisson observation model. As example forecasts we use the same simulation model but add error to either the growth rate or the mean of observations. We do this using the following steps,

- Define a grid of scenarios including the true values (zero growth rate or cumulative linear error).

```{targets toy_scenarios, tar_simple = TRUE}
expand_grid(
  init = c(100),
  r = c(0, 0.01, 0.1, -0.01, -0.1),
  r_error = c(0, 0.01, 0.025, 0.05, 0.1, -0.01, -0.025, -0.05, -0.1),
  additive_error = c(0, 1, 10, -1, -10)
) |> 
  mutate(id = 1:n())
```

- Simulate each scenario a 1000 times.

```{targets toy_simulations, tar_simple = TRUE}
toy_scenarios |>
  rowwise() |>
  mutate(y = list(
    simulate_exp_poisson(init = init, time = 4, growth = r + r_error,
                         additive_error = additive_error, sims = 1000)
    )
  ) |>
  unnest(cols = c(y))
```

- Filter for the "truth" scenario and keep only the first simulation.

```{targets toy_truth, tar_simple = TRUE}
toy_simulations |>
  filter(r_error == 0, additive_error == 0, sim == 1) |>
  select(init, r, time, true_value = obs)
```

- Filter for all other scenarios.

```{targets toy_forecasts, tar_simple = TRUE}
toy_simulations |>
  filter(!(r_error == 0 & additive_error == 0)) |>
  select(init, r, r_error, additive_error, time, sample = sim,
         prediction = obs) |>
  left_join(toy_truth, by = c("init", "r", "time"))
```

- Score example forecasts and summarise by growth rate, growth rate error, additive error, and forecast horizon.

```{targets, toy_scores, tar_simple = TRUE}
toy_forecasts |>
  eval_forecasts(
    summarise_by = c("init", "r", "r_error", "additive_error", "time"),
    metrics = c("crps")
  )
```

- Repeat the scoring process but on the log scale.

```{targets, toy_log_scores, tar_simple = TRUE}
toy_forecasts |>
  mutate(across(.cols = c("true_value", "prediction"), ~ log(.x))) |>
  eval_forecasts(
    summarise_by = c("init", "r", "r_error", "additive_error", "time"),
    metrics = c("crps")
  )
```

- Join log and natural scale scores.

```{targets, toy_all_scores, tar_simple = TRUE}
toy_log_scores |>
  mutate(scale = "log") |>
  bind_rows(
    toy_scores |>
      mutate(scale = "natural")
  )
```

- Plot natural scale scores by horizon.

```{targets plot_toy_scores_by_horizon, tar_simple = TRUE}
toy_scores |>
  filter(time <= 4) |>
  plot_scores_by_horizon()
```

- Plot log scale scores by horizon.

```{targets plot_toy_log_scores_by_horizon, tar_simple = TRUE}
toy_log_scores |>
  filter(time <= 4) |>
  plot_scores_by_horizon()
```

- Plot log and natural scale scores for the 4 week horizon.

```{targets plot_toy_scores_by_scale, tar_simple = TRUE}
toy_all_scores |>
  filter(time == 4) |>
  plot_scores_by_scale()
```

- Save output for external use.

```{targets save_toy_output}
list(
  tar_target(
    toy_scores_csv, 
    save_csv(toy_all_scores, here("output/data/toy_all_scores.csv")),
    format = "file"
  ),
  tar_target(
    toy_scores_by_scale_file, 
    save_plot(plot_toy_scores_by_scale, 
              here("output", "figures", "toy_scores_by_scale.png"),
    format = "file")
            ),
  tar_target(
    toy_scores_by_horizon_file, 
    save_plot(plot_toy_scores_by_horizon, 
              here("output", "figures", "toy_scores_by_horizon.png"),
    format = "file")
            ),
  tar_target(
    toy_log_scores_by_horizon_file, 
    save_plot(plot_toy_log_scores_by_horizon, 
              here("output", "figures", "toy_log_scores_by_horizon.png"),
    format = "file")
            )
)
```

# Pipeline

If all`{targets}` were run in non-interactive mode, then the pipeline is ready to run using the following,

```{r, eval = FALSE}
tar_make()
```

The complete pipeline can be visualised using,

```{r, eval = FALSE}
tar_visnetwork()
```