### 2021-09-28 _targets.Rmd
add oforecast plotting

---
title: "Supplementary information"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Setup

Set up the workflow pipeline and options. We first load the `targets` package and remove the potentially outdated workflow.

```{r}
library(targets)
library(tarchetypes)
library(tibble)
library(tidyr)
library(dplyr)
library(scoringutils)
library(ggplot2)
library(purrr)
library(here)
tar_unscript()
```

We now define shared global options across our workflow and load R functions from the `R` folder.

```{targets globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
library(purrr)
library(here)
library(tibble)
library(tarchetypes)
functions <- list.files(here("R"), full.names = TRUE)
walk(functions, source)
tar_option_set(
  packages = c("tibble", "tidyr", "dplyr", "scoringutils", "ggplot2", "purrr",
               "readr"),
  error = "workspace"
)
```

# Toy example

As a toy example we simulate a grid of data with varying exponential growth rates and a Poisson observation model. As example forecasts we use the same simulation model but add error to either the growth rate or the mean of observations. We do this using the following steps,

- Define a set of growth rates to explore.

```{targets toy_globals, tar_globals = TRUE}
toy_growth_rates <- tibble(r = c(0, 0.01, 0.1, -0.01, -0.1))
```

- Define a grid of scenarios including the true values (zero growth rate and cumulative linear error) for this set of growth rates.

```{targets toy_scenarios, tar_simple = TRUE}
expand_grid(
  init = c(100),
  r = toy_growth_rates$r,
  r_error = c(0, 0.01, 0.025, 0.05, 0.1, -0.01, -0.025, -0.05, -0.1),
  additive_error = c(0, 1, 10, -1, -10)
) |> 
  mutate(id = 1:n())
```

- Simulate each scenario a 1000 times.

```{targets toy_simulations, tar_simple = TRUE}
toy_scenarios |>
  rowwise() |>
  mutate(y = list(
    simulate_exp_poisson(init = init, time = 4, growth = r + r_error,
                         additive_error = additive_error, sims = 1000)
    )
  ) |>
  unnest(cols = c(y))
```

- Filter for the "truth" scenario and keep only the first simulation.

```{targets toy_truth, tar_simple = TRUE}
toy_simulations |>
  filter(r_error == 0, additive_error == 0, sim == 1) |>
  select(init, r, time, true_value = obs)
```

- Filter for all other scenarios.

```{targets toy_forecasts, tar_simple = TRUE}
toy_simulations |>
  select(init, r, r_error, additive_error, time, sample = sim,
         prediction = obs) |>
  left_join(toy_truth, by = c("init", "r", "time"))
```

- Plot simulated observations and forecasts for each growth rate stratified by assumed growth rate and additive error.

```{targets plot_toy_forecasts, tar_interactive = FALSE}
tar_map(
  toy_growth_rates,
  tar_target(
    toy_forecasts_plot,
    plot_toy_forecasts(toy_forecasts, growth = r[[1]],
                       alpha = 0.2, samples = 10)
  ),
  tar_target(
    plot_toy_forecasts_file,
    save_plot(
      toy_forecasts_plot, 
      here("output", "figures", 
           paste0("plot_toy_forecasts_", r[[1]], ".png"))
    ),
    format = "file")
)
```

- Score example forecasts and summarise by growth rate, growth rate error, additive error, and forecast horizon.

```{targets, toy_scores, tar_simple = TRUE}
toy_forecasts |>
  eval_forecasts(
    summarise_by = c("init", "r", "r_error", "additive_error", "time"),
    metrics = c("crps")
  )
```

- Repeat the scoring process but on the log scale.

```{targets, toy_log_scores, tar_simple = TRUE}
toy_forecasts |>
  mutate(across(.cols = c("true_value", "prediction"), ~ log(.x))) |>
  eval_forecasts(
    summarise_by = c("init", "r", "r_error", "additive_error", "time"),
    metrics = c("crps")
  )
```

- Join log and natural scale scores.

```{targets, toy_all_scores, tar_simple = TRUE}
toy_log_scores |>
  mutate(scale = "log") |>
  bind_rows(
    toy_scores |>
      mutate(scale = "natural")
  )
```

- Plot natural scale scores by horizon.

```{targets toy_scores_by_horizon_plot, tar_simple = TRUE}
toy_scores |>
  filter(time <= 4) |>
  plot_toy_scores_by_horizon()
```

- Plot log scale scores by horizon.

```{targets toy_log_scores_by_horizon_plot, tar_simple = TRUE}
toy_log_scores |>
  filter(time <= 4) |>
  plot_toy_scores_by_horizon()
```

- Plot log and natural scale scores for the 4 week horizon.

```{targets toy_scores_by_scale_plot, tar_simple = TRUE}
toy_all_scores |>
  filter(time == 4) |>
  plot_toy_scores_by_scale()
```

- Save output for external use.

```{targets save_toy_output}
list(
  tar_target(
    toy_scores_csv, 
    save_csv(toy_all_scores, here("output/data/toy_all_scores.csv")),
    format = "file"
  ),
  tar_target(
    toy_scores_by_scale_file, 
    save_plot(toy_scores_by_scale_plot, 
              here("output", "figures", "toy_scores_by_scale.png"),
    format = "file")
            ),
  tar_target(
    toy_scores_by_horizon_file, 
    save_plot(toy_scores_by_horizon_plot, 
              here("output", "figures", "toy_scores_by_horizon.png"),
    format = "file")
            ),
  tar_target(
    toy_log_scores_by_horizon_file, 
    save_plot(toy_log_scores_by_horizon_plot, 
              here("output", "figures", "toy_log_scores_by_horizon.png"),
    format = "file")
            )
)
```

# Pipeline

If all`{targets}` were run in non-interactive mode, then the pipeline is ready to run using the following,

```{r, eval = FALSE}
tar_make()
```

The complete pipeline can be visualised using,

```{r, eval = FALSE}
tar_visnetwork()
```