### 2021-10-25 analysis/Analysis.Rmd
update plots

---
title: "Log or not"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE, 
                      warning = FALSE)

library(scoringutils)
library(purrr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(here)

n <- 100
quantiles <- c(0.01, 0.025, seq(0.05, 0.95, by = 0.05), 0.975, 0.99)
rerun_simulation <- FALSE

```

Define a scenario grid with two data-generating distributions and two corresponding predictive distributions. For the predictive distributions one parameter is set equal to the true parameter, while the other is varied either in a multiplicative or an additive way.

```{r}
truth_mean_equal <- c(FALSE, TRUE)
truth_var_equal <- c(FALSE, TRUE)
error_var <- c("mu", "sigma")
error_type <- c("multiplicative", "additive")

scenarios <- expand_grid(truth_mean_equal, 
                         truth_var_equal, 
                         error_var, 
                         error_type)

grid_sigma_mult <- seq(0.5, 2, by = 0.1)
grid_mu_mult <- seq(0.5, 2, by = 0.1)
grid_sigma_add <- seq(-5, 10, by = 1)
grid_mu_add <- seq(-500, 1000, by = 100)

grids <- tibble(error_var = 
                  c("sigma", "mu", "sigma", "mu"),  
                error_type = 
                  c("multiplicative", "multiplicative", 
                    "additive", "additive"), 
                grid = list(grid_sigma_mult, 
                            grid_mu_mult, 
                            grid_sigma_add, 
                            grid_mu_add))

sigma <- 10
mu <- 1000

scenarios <- scenarios |>
  inner_join(grids) |>
  mutate(name = paste0(error_type, "_error_", error_var,  
                       ifelse(truth_mean_equal, "_equal_mean_", "_unequal_mean_"), 
                       ifelse(truth_var_equal, "equal_var", "unequal_var")), 
         title = paste(error_type, "error of", error_var, "for two variables with", 
                       ifelse(truth_mean_equal, "equal", "unequal"), 
                       "mean and", 
                       ifelse(truth_var_equal, "equal", "unequal"), "variance") |>
           str_to_sentence())

scenarios |>
  select(truth_mean_equal, truth_var_equal, error_var, error_type)

```

```{r, eval = rerun_simulation}
# helper functions
# ----------------

# function to generate a quantile forecast with a given error
create_forecast <- function(grid, sigma_true, mu_true,
                            var = "sigma", type = "multiplicative") {

  mu <- function(.x, v = var, t = type) {
    if (v == "mu") {
      if (t == "multiplicative") {
        return(mu_true * .x)
      } else if (t == "additive") {
        return(mu_true + .x)
      }
    } else if (var == "sigma") {
      return(mu_true)
    }
  }

  sigma <- function(.x, v = var, t = type) {
    if (var == "sigma") {
      if (t == "multiplicative") {
        return(sigma_true * .x)
      } else if (t == "additive") {
        return(sigma_true + .x)
      }
    } else if (var == "mu") {
      return(sigma_true)
    }
  }

  map_dfr(.x = grid, 
      .f = function(.x) {
        data.frame(prediction = 
                     qnorm(p = quantiles, 
                           mean = mu(.x), 
                           sd = sigma(.x)), 
                   quantile = quantiles, 
                   sigma_hat = sigma(.x), 
                   sigma_true = sigma_true,
                   mu_hat = mu(.x), 
                   mu_true = mu_true,
                   error = .x, 
                   error_type = type, 
                   error_var = var)
      }
  ) |>
  rowwise() |>
  mutate(n = list(1:n)) |>
  ungroup() |>
  unnest(cols = c(n))
}

# helper function to log predictions and true values
create_log_data <- function(data) {
  data |>
  mutate(prediction = log(prediction + 1), 
         true_value = log(true_value + 1))
}

# helper function to score forecasts
score_forecasts <- function(data) {
  eval_forecasts(data = data, 
                 metrics = c("interval_score"),
                 summarise_by = c("mu_true", "sigma_true", "error", 
                                  "error_var", "error_type")) 
}

# helper function to join the score data 
join_scores <- function(scores_1, scores_2, scale = "natural") {
  full_join(scores_1 |>
             select(error, sigma_true, mu_true, error_var, error_type, interval_score), 
           scores_2 |>
             select(error, sigma_true, mu_true, error_var, error_type, interval_score)
           ) |>
    mutate(scale = scale)
}

# helper function to wrap a long title
wrap_title <- function(x, ...) {
  paste(strwrap(x, ...), collapse = "\n")
}

# helper function to plot scores
plot_scores <- function(scores) {
  scores |>
    mutate("True parameters" = 
             paste0("mean: ", mu_true, ", sd: ", sigma_true)) |>
    group_by(`True parameters`, error, scale) |>
    summarise(interval_score = mean(interval_score)) |>
    group_by(scale) |>
    mutate(multplicative_score = interval_score / min(interval_score)) |>
    ggplot(aes(x = error, y = multplicative_score, 
               colour = `True parameters`, group = `True parameters`)) + 
    geom_line() + 
    scale_color_brewer(palette = "Set1", name = "True parameters: ") + 
    scale_y_continuous(limits = c(1, NA), expand = c(1, NA)) + 
    expand_limits(y = c(1, NA)) + 
    facet_wrap(~ scale, scale = "free") + 
    theme_minimal() + 
    theme(legend.position = "bottom") 
}

```

```{r eval = rerun_simulation}

results <- list()

for (i in 1:nrow(scenarios)) {
  equal_mean <- scenarios[i, ]$truth_mean_equal
  equal_var <- scenarios[i, ]$truth_var_equal
  grid <- unlist(scenarios[i, ]$grid)
  error_var <- scenarios[i, ]$error_var
  error_type <- scenarios[i, ]$error_type
  scenario_name <- scenarios[i, ]$name
  scenario_title <- scenarios[i, ]$title

  sigma_true <- c(sigma, sigma + (as.numeric(!equal_var) * 90))
  mu_true <- c(mu, mu + as.numeric(!equal_mean) * 1000)

  true_values_1 <- 
    data.frame(
      true_value = rnorm(n = n, mean = mu_true[1], sd = sigma_true[1]),
      n = 1:n,
      sigma_true = sigma_true[1]
    )

  true_values_2 <- 
    data.frame(
      true_value = rnorm(n = n, mean = mu_true[2], sd = sigma_true[2]),
      n = 1:n,
      sigma_true = sigma_true[2]
    )

  forecasts_1 <- create_forecast(grid, 
                                 sigma_true = sigma_true[1],
                                 mu_true = mu_true[1], 
                                 var = error_var, 
                                 type = error_type)

  forecasts_2 <- create_forecast(grid, 
                                 sigma_true = sigma_true[2],
                                 mu_true = mu_true[2], 
                                 var = error_var, 
                                 type = error_type)

  data_1 <- inner_join(forecasts_1, true_values_1)
  data_2 <- inner_join(forecasts_2, true_values_2)
  log_data_1 <- create_log_data(data_1)
  log_data_2 <- create_log_data(data_2)

  scores_1 <- score_forecasts(data_1)
  scores_2 <- score_forecasts(data_2)
  log_scores_1 <- score_forecasts(log_data_1)
  log_scores_2 <- score_forecasts(log_data_2)

  joined_scores <- join_scores(scores_1, scores_2)
  joined_log_scores <- join_scores(log_scores_1, log_scores_2, scale = "log")
  combined_scores <- rbind(joined_scores, joined_log_scores) |>
    mutate(equal_mean = equal_mean, 
           equal_var = equal_var)

  # store all results
  results[[scenario_name]] <- combined_scores

  plot_scores(combined_scores) + 
    ggtitle(wrap_title(scenario_title, width = 60)) + 
    labs(y = "Multiplicative change in Interval Score", 
         x = paste(error_type, "error on", error_var) |> 
           str_to_sentence())

  ggsave(here("output", "figures", paste0(scenario_name, ".png")))
}

```

```{r eval=rerun_simulation}
results

res <- bind_rows(results)

plot_multiple <- function(res, equal_means, equal_vars) {
  res <- res |>
    filter(equal_mean == equal_means, 
           equal_var == equal_vars) |>
    mutate("True parameters" = 
             paste0("mean: ", mu_true, ", sd: ", sigma_true)) |>
    group_by(`True parameters`, error, error_type, error_var, scale) |>
    summarise(interval_score = mean(interval_score)) |>
    group_by(scale, error_var, error_type) |>
    mutate(multplicative_score = interval_score / min(interval_score), 
           error_title = str_to_title(
             paste(error_type, "on", error_var)))

  res |> 
    ggplot(aes(x = error, y = multplicative_score, 
               colour = `True parameters`, group = `True parameters`)) + 
    geom_line() + 
    scale_color_brewer(palette = "Set1", name = "True parameters: ") + 
    # scale_y_continuous(limits = c(1, NA), expand = c(0, NA)) + 
    expand_limits(y = c(1, NA)) + 
    facet_wrap( ~ error_title + scale, scale = "free", nrow = 2) + 
    theme_minimal() + 
    theme(legend.position = "bottom") +
    labs(y = "Multiplicative change in Interval Score", 
         x = "Error") + 
    geom_rect(data = subset(res, error_var == "sigma"), 
              fill = NA, colour = "grey80", xmin = -Inf,xmax = Inf,
              ymin = -Inf,ymax = Inf) + 
    geom_rect(data = subset(res, error_var == "mu"), 
              fill = NA, colour = "grey70", xmin = -Inf,xmax = Inf,
              ymin = -Inf,ymax = Inf) 
}

p1 <- plot_multiple(res, equal_means = TRUE, equal_vars = TRUE) 
ggsave(p1, filename = here("output", "effect_equal_mean_equal_var.png"))

p2 <- plot_multiple(res, equal_means = FALSE, equal_vars = TRUE) 
ggsave(p2, filename = here("output", "effect_unequal_mean_equal_var.png"))

p3 <- plot_multiple(res, equal_means = TRUE, equal_vars = FALSE) 
ggsave(p3, filename = here("output", "effect_equal_mean_unequal_var.png"))

p4 <- plot_multiple(res, equal_means = FALSE, equal_vars = FALSE) 
ggsave(p4, filename = here("output", "effect_unequal_mean_unequal_var.png"))

```

---

Effect of scoring on a log scale

```{r}
knitr::include_graphics(here("output", "effect_equal_mean_equal_var.png"))
```

---

Effect of scoring on a log scale for two variables with unequal means

```{r}
knitr::include_graphics(here("output", "effect_unequal_mean_equal_var.png"))
```

---

Effect of scoring on a log scale for two variables with unequal variances

```{r}
knitr::include_graphics( here("output", "effect_equal_mean_unequal_var.png"))
```

Effect of scoring on a log scale for two variables with unequal means and unequal variances

```{r}
knitr::include_graphics(here("output", "effect_unequal_mean_unequal_var.png"))
```

---

```{r}
i = 1
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 2
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 3
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 4
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 5
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 6
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 7
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 8
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 9
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))

```

What we learn from this plot:

- with a multiplicative error in the mean and the correct standard deviation, the standard deviation of the underlying data makes some difference, but not that much
- multiplicative errors in the mean get punished much more harshly when scoring on the natural scale compared to scoring on the log scale

---

```{r}
i = 10
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 11
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

What we learn from this plot:

- A perfect forecast that gets the mean as well as the variance right will incur higher scores if the true variance is higher. 
- getting the variance wrong (in multiplicative terms) is equally costly regardless of what the true variance is (curves look identical on a log scale plot, I checked that previously)
- Those two effects are exactly the same on the linear as on the log scale? Not sure I understand this

---

```{r}
i = 12
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 13
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 14
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 15
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```

---

```{r}
i = 16
file_name <- paste0(scenarios$name[i], ".png")
knitr::include_graphics(here("output", "figures", file_name))
```