### 2022-11-29 analysis/Analysis ECDC forecasts.Rmd
update plots

---
title: "Log or not - Analysis of ECDC forecasts"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE, 
                      warning = FALSE)

library(scoringutils)
library(purrr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(here)
library(data.table)
library(kableExtra)
library(performance)
library(tidytext)
library(grid)
library(ggpubr)

recompute_scores <- FALSE

# situations with negative values: France 2021-05-22, CZ 2021-08-07, ES 2021-03-06

# helper functions
# ------------------------------------------------------------------------------

make_table <- function(x, caption = "") {
  x %>%
    rename_all(gsub, pattern = "_", replacement = " ") %>%
    rename_all(str_to_sentence) %>%
    mutate_if(is.numeric, round, 2) %>%
    kable(caption = caption) %>%
    kable_styling()
}

# format plot scales
scale_fn <- function(x) {
  ifelse(x >= 1000000, 
         paste0(x / 1000000, "m"), 
         ifelse(x >= 1000,
                paste0(x / 1000, "k"),
                x))
}

label_fn <- function(x) {
  ifelse(x < 1000, 
         paste(x), 
         ifelse(x < 1e6, 
                paste0(x / 1000, "k"),
                ifelse(x < 1e9, 
                       paste0(x / 1e6, "m"), 
                       paste0(x / 1e9, "b"))
                )
         )
}

# load data 
# ------------------------------------------------------------------------------
hub_data <- rbindlist(list(
  fread(here("data", "full-data-european-forecast-hub-1.csv")), 
  fread(here("data", "full-data-european-forecast-hub-2.csv")),
  fread(here("data", "full-data-european-forecast-hub-3.csv"))
)) |>
  unique()

scores <- fread(here("output", "data", "all-scores-european-hub.csv"))
scores[, scale := factor(scale, levels = c("natural", "log"))]
data.table::setnames(scores, old = c("aem", "sharpness"), new = c("ae_median", "dispersion"))

scores_gr <- fread(here("output", "data", "all-scores-growth-rate-european-hub.csv"))
scores_gr[, scale := factor(scale, levels = c("natural", "log"))]

check <- hub_data |> 
  check_forecasts()

filtermodels <- check$unique_values |>
  filter(horizon >= 4, 
         target_type == 2,
         true_value > 900,
         quantile == 23, 
         forecast_date > 16,
         location == 32) |>
  pull(model)

remove <- hub_data |> 
  group_by(model, forecast_date, location, target_type) |>
  summarise(q = length(unique(quantile))) |>
  filter(q < 23) |>
  pull(model) |>
  unique()

filtermodels <- setdiff(filtermodels, remove)

scores <- scores[model %in% filtermodels]
scores[, type_and_scale := paste0(target_type, " - ", scale)]
scores[, type_and_scale := factor(type_and_scale, 
                                  levels = c("Cases - natural", "Deaths - natural",
                                             "Cases - log", "Deaths - log"))]

hub_data <- hub_data[model %in% filtermodels]
hub_data <- hub_data[forecast_date >= "2021-03-05"]
hub_data[, true_value := pmax(true_value, 0)]

```

```{r eval = recompute_scores}
# score forecasts
# ------------------------------------------------------------------------------

# helper function
score_forecasts <- function(data,
                            summarise_by = c("model", "location", 
                                             "target_end_date", "forecast_date",
                                             "horizon", "target_type"), 
                            scale = c("natural", "log")) {
  if (scale[1] == "log") {
    data <- data |>
      mutate(prediction = log(pmax(prediction, 0) + 1), 
             true_value = log(pmax(true_value, 0) + 1))
  }

  data |>
    eval_forecasts(summarise_by = summarise_by) |>
    mutate(scale = scale[1])
}

scores_natural <- score_forecasts(hub_data)
scores_log <- score_forecasts(hub_data, scale = "log")

scores_gr_natural <- score_forecasts(growth_data)
scores_gr_log <- score_forecasts(growth_data, scale = "log")

scores <- bind_rows(scores_natural, scores_log)
fwrite(scores, here("output", "data", "all-scores-european-hub.csv"))

scores_gr <- bind_rows(scores_gr_natural, scores_gr_log)
fwrite(scores_gr, here("output", "data", "all-scores-growth-rate-european-hub.csv"))
```

## =============================================================================
## FIGURE 1
# illustration-effect-log-score.R

## =============================================================================
## FIGURE 2
# Simulation poisson-mean-sd.R

## =============================================================================
## FIGURE 3
# example-log-first-Johannes.R

## =============================================================================
## FIGURE 4
## Mean vs. sd of observed values + boxplots for distribution across target type + location

```{r}
# average number of cases and deaths

# plot with means across locations ---------------------------------------------
label_fn_within <- function(x) {
  x <- gsub(pattern = "___Deaths", "", x)
  x <- gsub(pattern = "___Cases", "", x)
  paste(x)
}

mean_observations <- hub_data |>
  dplyr::select(location, target_type, target_end_date, true_value) |>
  unique() |>
  group_by(target_type, location) |>
  summarise(mean_value = mean(true_value), 
            sd_value = sd(true_value), 
            var_value = var(true_value))

plot_means_obs <- mean_observations |>
  ggplot(aes(y = mean_value, x = reorder_within(location, -mean_value, target_type))) +   geom_bar(stat = "identity", fill = "grey50") + 
  facet_wrap(~ target_type, scales = "free") + 
  theme_scoringutils() + 
  scale_y_continuous(labels = label_fn) + 
  scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) + 
  labs(y = "Mean observed value", x = "") +
  theme(axis.title.x = element_blank())

box_plot_obs <- hub_data |>
  mutate(true_value = pmax(true_value, 1)) |>
  dplyr::select(location, target_type, target_end_date, true_value) |>
  unique() |>
  ggplot(aes(y = true_value, x = target_type)) + 
  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
  geom_boxplot(alpha = 0.5) + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn, trans = "log10") + 
  labs(y = "Observations", x = "")  +
  theme(axis.title.x = element_blank())

plot_df <- scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(target_type, location, scale, type_and_scale) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") |> 
  full_join(mean_observations)

plot_mean_scores <- plot_df |>
  ggplot(aes(y = interval_score,
             x = reorder_within(location, -mean_value, target_type))) + 
  geom_bar(stat = "identity") + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ type_and_scale, scale = "free") + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn) + 
  scale_x_discrete(guide = guide_axis(n.dodge=2), labels = label_fn_within) +
  labs(y = "Mean interval score", x = "Loaction")

box_plot_scores <- scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  ggplot(aes(y = interval_score, x = target_type)) + 
  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
  geom_boxplot(alpha = 0.5) + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ scale, scale = "free", nrow = 2) + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn, trans = "log10") + 
  labs(y = "Interval score", x = "Target type")

box_plot_horizon <- scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon <= 4) |>
  group_by(forecast_date, model, location, target_type, scale) |>
  mutate(interval_score = interval_score / interval_score[horizon == 1]) |>
  ggplot(aes(y = interval_score,
             x = as.factor(horizon), fill = target_type)) + 
  geom_hline(aes(yintercept = 1), linetype = "dashed", color = "grey40") +
  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) +
  geom_boxplot(alpha = 0.5, position = "dodge2") + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ scale, nrow = 1) + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn, trans = "log10") +
  labs(y = "Rel. change in WIS", x = "Forecast horizon (weeks)")

layout <- "
AB
CD
CD
EF"

plot_means_obs  + box_plot_obs + plot_mean_scores + box_plot_scores + box_plot_horizon + plot_spacer() +
  plot_layout(heights = c(1, 2, 1), 
              widths = c(3, 1)) +
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A")

ggsave("output/figures/HUB-mean-obs-location.png", width = 10, height = 10)

# This uses the ensemble, but alternatively, we could think about just summarising across all models for these plots? 
```

## =============================================================================
## Figure 5 - Correlations between everything
```{r}
wis_vs_total <- function(scores, summary_fct = mean, fulldata = hub_data) {

  mean_values <- 
    fulldata |> 
    filter(!is.na(prediction)) |>
    group_by(location, target_type, target_end_date) |>
    summarise(true_value = unique(true_value)) |>
    group_by(location, target_type) |>
    summarise(mean_obs = mean(true_value))

  plot_df <- scores |> 
    filter(model == "EuroCOVIDhub-ensemble", 
           horizon == 2) |>
    group_by(location, target_type, scale, type_and_scale) |>
    summarise(wis = summary_fct(interval_score), 
              .groups = "drop_last") |> 
    inner_join(mean_values)

  plot_df |>
    group_by(target_type, scale) |>
    summarise(correlation = cor(wis, mean_obs)) |> 
    print()

  plot_df |>
    ggplot(aes(y = wis, x = mean_obs)) + 
    geom_point() + 
    facet_wrap(~ type_and_scale, scale = "free") + 
    theme_scoringutils() + 
    theme(legend.position = "none") + 
    scale_x_continuous(labels = scale_fn) +
    scale_y_continuous(labels = scale_fn) +
    labs(y = "REPLACE", x = "Mean observed value")
}

scatter_wis_obs <- wis_vs_total(scores) + 
  labs(y = "Mean WIS") + 
  scale_x_continuous(labels = scale_fn, trans = "log10") +
  scale_y_continuous(labels = scale_fn, trans = "log10") + 
  ggpubr::stat_cor(method = "spearman", p.accuracy = 0.001, 
                   cor.coef.name = "rho", size = 3)

df <- scores |>
  filter(horizon < 5) |>
  select(-c(coverage_deviation, bias, ae_median, dispersion, underprediction, overprediction, type_and_scale)) |>
  pivot_wider(names_from = scale, values_from = interval_score) |> 
  group_by(horizon, target_type, forecast_date, target_end_date, location) |>
  summarise(cor = cor(natural, log), 
            spearman = cor(natural, log, method = "spearman")) |>
  group_by(horizon, target_type) |>
  summarise(cor = mean(cor, na.rm = TRUE), 
            spearman = mean(cor, na.rm = TRUE)) |>
  rename(`Target type` = target_type)

p_cor_scores <- df |>
  ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
  scale_fill_brewer(palette = "Set1", name = "Target type") + 
  scale_color_brewer(palette = "Set1", name = "Target type") +
  geom_histogram(stat = "identity", position = "dodge", alpha = 0.4) +
  theme_scoringutils() + 
  coord_cartesian(ylim = c(0, 1)) +
  labs(y = "Avg. rank cor(scores)", x = "Forecast horizon") + 
  theme(axis.title.y = element_text(size=8.5))

# Correlation between pairwise comparisons
correlation_rel_skill <- scores |>
  filter(horizon < 5) |>
  summarise_scores(
    by = c("model", "scale", "horizon", "target_type"), 
    relative_skill = TRUE) |>
  select(model, horizon, target_type, relative_skill, scale) |>
  pivot_wider(values_from = relative_skill, names_from = scale) |>
  group_by(horizon, target_type) |>
  summarise(cor = cor(natural, log), 
            spearman = cor(natural, log, method = "spearman"))

p_cor_skill <- correlation_rel_skill |>
  rename(`Target type` = target_type) |>
  ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
  scale_fill_brewer(palette = "Set1", name = "Target type") + 
  scale_color_brewer(palette = "Set1", name = "Target type") + 
  geom_histogram(stat = "identity", position = "dodge", alpha = 0.4) +
  theme_scoringutils() + 
  coord_cartesian(ylim = c(0, 1)) +
  labs(y = "Cor(rel. skill)", x = "Forecast horizon")

(scatter_wis_obs) / (p_cor_scores + p_cor_skill) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") +
  plot_layout(widths = c(1, 1), heights = c(3, 1)) &
  theme(legend.position = "bottom")

ggsave("output/figures/HUB-correlations.png", width = 7, height = 7)

```

## =============================================================================
## Figure 6 - Example comparisons of forecasts on the log and the natural scale

```{r}
plot_pred_score <- function(hub_data, scores, model,
                            horizon = 2, type = "Cases", 
                            remove_x_text = TRUE) {
  forecast_model <- model
  h <- horizon
  filtered_hub <- hub_data |>
    filter(model == forecast_model, 
           horizon %in% h, 
           target_type == type)

  locations <- unique(filtered_hub$location)

  filtered_scores <- scores |>
    filter(model == forecast_model, 
           horizon %in% h, 
           location %in% locations,
           target_type == type)

  plot_natural  <- filtered_hub |>  
    mutate(id = paste(type, "- natural")) |>
    plot_predictions(x = "target_end_date") +
    scale_y_continuous(labels = label_fn)  +
    facet_wrap(~ id) +
    theme(legend.position = "bottom", 
          axis.text.x = element_blank(), 
          axis.title.x = element_blank()) +
    labs(y = "Predictions")

  plot_log <- filtered_hub |>
    mutate(prediction = log(prediction + 1), 
           true_value = log(true_value + 1)) |>
    mutate(id = paste(type, "- log")) |>
    plot_predictions(x = "target_end_date") +
    scale_y_continuous(labels = label_fn)  +
    facet_wrap(~ id) +
    theme(legend.position = "bottom", 
          axis.text.x = element_blank(), 
          axis.title.x = element_blank()) +
    labs(y = "Predictions")

  scores_natural <- filtered_scores |>
    filter(scale == "natural") |>
    summarise_scores("target_end_date") |>
    mutate(target_end_date = factor(target_end_date)) |>
    plot_wis(x = "target_end_date", flip = TRUE) +
    scale_x_continuous(labels = label_fn) +
    labs(y = "Target end date")

  scores_log <- filtered_scores |>
    filter(scale == "log") |>
    summarise_scores("target_end_date") |>
    mutate(target_end_date = factor(target_end_date)) |>
    plot_wis(x = "target_end_date", flip = TRUE) +
    labs(y = "Target end date")

  if (remove_x_text) {
    scores_natural <- scores_natural +
      theme(axis.text.x = element_blank(), 
            axis.title.x = element_blank())
    # scores_log <- scores_log +
    #   theme(axis.text.x = element_blank(), 
    #         axis.title.x = element_blank())
  }

  list(plot_natural, scores_natural, plot_log, scores_log )
}

put_plot_together <- function(modelname = "EuroCOVIDhub-ensemble", 
                              locationname = "DE") {

  plot_cases <- plot_pred_score(hub_data |> filter(location == locationname), 
                                  scores, modelname, 
                                  remove_x_text = TRUE, 
                                  type = "Cases")

  plot_deaths <- plot_pred_score(hub_data |> filter(location == locationname), 
                                     scores, modelname, 
                                     type = "Deaths")

  layout <- "
AE
BF
CG
DH"

  plot_cases[[1]] + plot_cases[[2]] + plot_cases[[3]] +
    plot_cases[[4]] + plot_deaths[[1]] +
    plot_deaths[[2]] + plot_deaths[[3]] + plot_deaths[[4]] +
    plot_annotation(tag_levels = "A") +
    plot_layout(guides = "collect", 
                design = layout) &
    theme(legend.position = "bottom")
}

put_plot_together("EuroCOVIDhub-ensemble")
ggsave(filename = "output/figures/HUB-model-comparison-ensemble.png", width = 10, height = 8.5)

put_plot_together("epiforecasts-EpiNow2")
ggsave(filename = "output/figures/HUB-model-comparison-epinow.png", width = 10, height = 8.5)

put_plot_together("EuroCOVIDhub-baseline", locationname = "DE")
ggsave(filename = "output/figures/HUB-model-comparison-baseline.png", width = 10, height = 8.5)

```

## =============================================================================
## Figure 7 - Look at pairwise comparisons and changes in rankings (plus Appendix figures)

```{r}

ranking_figure <- function(target = "Cases") {
  summarised_pairwise <- scores |>
  filter(target_type == target, 
         horizon == 2) |>
  summarise_scores(by = c("model", "scale"), relative_skill = TRUE)

ranking_log <- 
  summarised_pairwise |>
  filter(scale == "log") |>
  arrange(-relative_skill) |>
  pull(model)

plot_rel_skill_natural <- summarised_pairwise |>
  filter(scale == "natural") |>
  ggplot(aes(y = reorder(model, -relative_skill), x = relative_skill)) +
  geom_bar(stat = "identity") +
  theme_scoringutils() +
  facet_wrap(~scale) +
  labs(y = "Model", x = "Relative skill")

plot_rel_skill_log <- summarised_pairwise |>
  filter(scale == "log") |>
  ggplot(aes(y = reorder(model, -relative_skill), x = relative_skill)) +
  geom_bar(stat = "identity") +
  theme_scoringutils() +
  facet_wrap(~scale) +
  labs(y = "", x = "Relative skill") +
  theme(axis.title.y = element_blank())

diffs <- summarised_pairwise |>
  select(model, relative_skill, scale) |>
  pivot_wider(names_from = scale, values_from = relative_skill) |>
  mutate(model = factor(model, levels = ranking_log), 
         difference = log - natural)

plot_diffs <- diffs |>
  mutate(title = "difference") |>
  ggplot(aes(y = model, x = difference)) +
  geom_bar(stat = "identity") +
  theme_scoringutils() +
  labs(y = "Model", x = "\u0394 Relative skill") +
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.line.y = element_blank()) +
  facet_wrap(~title)

plot_rel_wis <- scores |>
  filter(horizon == 2, 
         target_type == "Cases") |>
  mutate(model = factor(model, levels = ranking_log)) |>
  summarise_scores(by = c("model", "scale")) |>
  plot_wis(relative_contributions = TRUE) +
    theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.line.y = element_blank()) +
  scale_x_continuous(labels = label_fn) +
  facet_wrap(~scale, scales = "free_x")

plot_ranking_change <- summarised_pairwise |>
  select(model, scale, relative_skill) |>
  group_by(scale) |>
  mutate(rank = rank(-relative_skill), 
         x = ifelse(scale == "natural", 0.4, 2.8), 
         x_arrow = ifelse(scale == "natural", 1.7, 2.7)
         ) |>
  group_by(model) |>
  mutate(color = ifelse(diff(relative_skill) > 0, "deteriorated", "not deteriorated")) |>
  ggplot(aes(y = rank, x = x, group = model, label = model)) +
    geom_path(
      aes(x=x_arrow, color = color), 
      arrow = arrow(length = unit(0.09,"npc")), 
      lineend = "round", linejoin = "mitre",
      size=1, show.legend = FALSE) +
  # geom_text(size=3.3, hjust = 0) +
  coord_cartesian(#xlim = c(0.5, 4), 
                  ylim = c(0.75, 8.25)) +
  theme_scoringutils() +
  theme(axis.title = element_blank(),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank(), 
        legend.position="none")

plot_rel_wis + plot_ranking_change

# layout <- "
# AAA
# BCD"

return(
  list(
    plot_ranking_change, plot_rel_skill_natural, plot_rel_skill_log, plot_diffs, plot_rel_wis
  )
)
}

ranking_cases <- ranking_figure(target = "Cases")
ranking_deaths <- ranking_figure(target = "Deaths")

layout <- "
ABCDE
FGHIJ"

ranking_cases[[2]] + ranking_cases[[1]] + 
ranking_cases[[3]] + ranking_cases[[4]] + ranking_cases[[5]] +
ranking_deaths[[2]] + ranking_deaths[[1]] + 
ranking_deaths[[3]] + ranking_deaths[[4]] + ranking_deaths[[5]] +
  plot_annotation(tag_levels = "A") +
  plot_layout(design = layout, 
              guides = "collect", 
              widths = c(1, 1, 1, 1, 2.5)) &
  theme(legend.position = "bottom")

ggsave("output/figures/HUB-pairwise-comparisons.png", width = 10.5, height = 6)

# summarised_pairwise |>
#   select(model, scale, relative_skill) |>
#   pivot_wider(names_from = scale, values_from = relative_skill) |>
#   ggplot(aes(y = log, x = natural)) +
#   geom_point() + 
#   # geom_abline(slope = 1) +
#   theme_scoringutils()

```

## =============================================================================
## Summary numbers for the text + Figure Available Forecasts fo Appendix

```{r}
hub_data |>
  summarise(min_date = min(forecast_date), 
            max_date = max(forecast_date)) |>
  mutate(week_diff = (max_date - min_date - 1) / 7)

hub_data$location |> 
  unique() |>
  length()

hub_data |> 
  filter(horizon == 2) |>
  avail_forecasts(by = c("target_end_date", "target_type", "model")) |>
  mutate(forecast_date = as.Date(target_end_date) - 12) |>
  plot_avail_forecasts(show_numbers = FALSE) +
  facet_grid(~ target_type) + 
  labs(y = "Model", x = "Forecast date")

ggsave(filename = "output/figures/number-avail-forecasts.png", height = 4, width = 10)
```

## =============================================================================
## Summary table for Appendix

# Table with relevant values

```{r}
summary <- hub_data |>
  filter(horizon == 2) |>
  mutate(natural = true_value, 
         log = log(pmax(0, true_value) +1)) |>
  select(location, target_type, target_end_date, natural, log) |>
  pivot_longer(cols = c(natural, log), values_to = "true_value", names_to = "scale") |>
  unique() |>
  group_by(target_type, scale) |>
  summarise(mean = mean(true_value), 
            sd = sd(true_value), 
            var = var(true_value))

table_obs <- summary |>
  pivot_longer(cols = c(mean, sd, var), names_to = "measure") |>
  pivot_wider(values_from = value, names_from = scale) |>
  select(target_type, measure, natural, log) |>
  mutate(quantity = "Observations", 
         natural = round(natural), 
         log = round(log, 2)) |>
  select(target_type, quantity, measure, natural, log)

table_scores <- scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(target_type, scale) |>
  summarise(mean = mean(interval_score), 
            sd = sd(interval_score), 
            .groups = "drop_last") |>
  pivot_longer(cols = c(mean, sd), names_to = "measure") |>
  pivot_wider(values_from = value, names_from = scale) |>
 mutate(quantity = "WIS", 
         natural = round(natural), 
         log = round(log, 2)) |>
  select(target_type, quantity, measure, natural, log)

table <- rbind(table_obs, 
               table_scores) |>
  arrange(quantity, target_type, measure)

table |>
  kable(format = "latex", booktabs = TRUE,
        align = c("l", "l", "l", rep("c", 2)),
        # col.names = c(" ", names(scores)[-(1:2)])) %>%
  ) #|>
  # collapse_rows(columns = 1) %>%
  # kable_styling(latex_options = c("scale_down",
  #                                 "hold_position")) %>%
  # column_spec(1, background = "white") %>%
  # pack_rows("Observations", 1, 6, indent = FALSE,
  #           hline_after = TRUE) %>%
  # pack_rows("WIS", 7, 10, indent = FALSE,
  #           hline_after = TRUE)

```

## Scores across forecast targets

This uses the Hub ensemble

```{r}
scores |>
  dplyr::select(location, target_type, scale, target_end_date, interval_score) |>
  unique() |>
  group_by(target_type, scale) |>
  summarise(mean_value = mean(interval_score), 
            sd_value = sd(interval_score), 
            var_value = var(interval_score))

ggsave("output/figures/HUB-average-scores.png", width = 7, height = 2.5)
```

Mean scores: 
```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(scale, target_type) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") |>
  make_table()
```

## Correlation between scores

```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(target_type, location, scale) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") %>%
  pivot_wider(values_from = interval_score, names_from = scale) %>%
  group_by(target_type) %>%
  summarise(correlation = cor(natural, log))
```

# Correlations for ranks, scores and pairwise comparisons

```{r}

# 1. observations natural vs. log --> pearson
# 2. correlation scores natural vs. log (single forecasts) --> spearman

# Table
# correlation between values
cor_data <- hub_data |>
  select(location, target_end_date, target_type, true_value) |>
  unique() |>
  mutate(log_obs = log(pmax(true_value, 0) + 1)) |>
  group_by(target_type) |>
  summarise(cor = cor(true_value, log_obs))

# # correlation between scores
# # not really meaningful: if we do ranks it's just going to be ranks over everything which doesn't make sense
# # if we do individual scores than it's a bit unclear what the "correct" correlation should be
# cor_scores <- scores |>
#   filter(horizon <=4) |>
#   select(interval_score, scale, forecast_date, model, horizon, location, target_type) |>
#   pivot_wider(names_from = scale, values_from = interval_score) |>
#   group_by(horizon, target_type) |>
#   summarise(cor = cor(natural, log), 
#             spearman = cor(natural, log, method = "spearman"))

# Correlation between ranks when going from natural to log
correlation_ranks <- rankings_wide |>
  filter(horizon < 5) |>
  group_by(horizon, target_type) |>
  summarise(cor = cor(natural, log), 
            spearman = cor(natural, log, method = "spearman")) |>
  rename(`Target type` = target_type)

# we could also do spearman per target and then plot an average of the spearman correlation.

# Correlation between pairwise comparisons
correlation_rel_skill <- scores |>
  filter(horizon < 5) |>
  summarise_scores(
    by = c("model", "scale", "horizon", "target_type"), 
    relative_skill = TRUE) |>
  select(model, horizon, target_type, relative_skill, scale) |>
  pivot_wider(values_from = relative_skill, names_from = scale) |>
  group_by(horizon, target_type) |>
  summarise(cor = cor(natural, log), 
            spearman = cor(natural, log, method = "spearman"))

p_cor_ranks <- correlation_ranks |>
  ggplot(aes(x = horizon, y = cor, 
             fill = `Target type`, color = `Target type`)) + 
  geom_histogram(stat = "identity", position = "dodge") +
  theme_scoringutils() + 
  coord_cartesian(ylim = c(0, 1)) +
  labs(y = "Corr. ranks per forecast", x = "Forecast horizon")

# p_cor_scores <- cor_scores |> 
#   mutate(`Target type` = target_type) |>
#   ggplot(aes(x = horizon, y = cor, 
#              fill = `Target type`, color = `Target type`)) + 
#   geom_histogram(stat = "identity", position = "dodge") +
#   theme_scoringutils() + 
#   coord_cartesian(ylim = c(0, 1)) +
#   labs(y = "Spearman corr. scores single forecasts", x = "Forecast horizon")

p_cor_skill <- correlation_rel_skill |>
  rename(`Target type` = target_type) |>
  ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
  geom_histogram(stat = "identity", position = "dodge") +
  theme_scoringutils() + 
  coord_cartesian(ylim = c(0, 1)) +
  labs(y = "Correlation rel. skill", x = "Forecast horizon")

combined <- scores[, c("model", "location", "target_end_date", "scale", "target_type", "interval_score", "horizon")] |>
  full_join(hub_data[, c("location", "target_end_date", "target_type", "model", "true_value")]) |>
  mutate(true_value = ifelse(scale == "natural", true_value, log(true_value + 1)))

combined |>
  filter(horizon <= 4) |>
  rename(`Target type` = target_type) |>
  group_by(horizon, scale, `Target type`) |>
  summarise(cor = cor(interval_score, true_value)) |>
    ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
  geom_histogram(stat = "identity", position = "dodge") +
  theme_scoringutils() + 
  facet_wrap(~ scale) +
  expand_limits(ylim = c(0, 1)) +
  labs(y = "Correlation score and observation", x = "Forecast horizon")

plot_cors <- p_cor_ranks  + p_cor_skill +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "bottom")

ggsave(filename = "output/figures/HUB-cors.png", width = 7, height = 3, 
       plot = plot_cors)

```

```{r}
scores |>
  summarise_scores(c("model", "target_type", "scale")) |>
  summarise_scores(c("model", "target_type", "scale"), fun = signif, digits = 2) |>
  plot_score_table(by = c("target_type", "scale")) +
  facet_wrap(target_type ~ scale)
```

```{r}

split_data <- split(scores, by = c("type_and_scale"))
cors <- lapply(split_data, FUN = correlation, 
               metrics = c("interval_score", 
                            "bias", "overprediction", 
                            "underprediction",
                            "dispersion"))

# scoringutils dev stuff
# split_data <- split(scores, by = c("target_type", "scale"))
# list_name_as_col <- function(list, by = NULL) {
#   names <- names(list)
#   print(names(list_element))
# }

plots <- lapply(cors, plot_correlation)

(plots[[1]] + plots[[2]]) /
  (plots[[3]] + plots[[4]])

correlation
```

```{r}
s <- hub_data |>
  filter(model %in% c("EuroCOVIDhub-ensemble", "epiforecasts-EpiNow2")) |>
  score()

s_summarised <- s |>
  summarise_scores(by = c("model", "range", "horizon"), metric = c("interval_score"))

s_summarised |>
  select(c(model, interval_score, range, horizon)) |>
  pivot_wider(names_from = model, values_from = interval_score) |>
  mutate(rel = `epiforecasts-EpiNow2` / `EuroCOVIDhub-ensemble`) |>
  filter(horizon == 4)
  ggplot(aes(y = rel, x = range, 
             color = factor(horizon), group = factor(horizon))) +
  geom_line()

s_summarised |>
  filter(horizon == 1) |>
  select(c(model, interval_score, range, horizon)) |>
  ggplot(aes(y = interval_score, x = range, 
             color = model, group = model)) +
  geom_line()

```

--- 
# Code parking lot

## Scores over time
restrict to a few locations, e.g. GB, FR, ES (highest average scores there)

```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble",
         location %in% c("GB", "ES", "DE"),
         horizon == 2) |>
  rename(Location = location) |>
  group_by(target_type, scale, Location) |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = forecast_date, colour = Location)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale, scale = "free_y") + 
  theme_scoringutils() + 
  labs(x = "Forecast date", y = "WIS relative to smallest observed value")

ggsave("output/figures/HUB-scores-over-time.png", width = 7, height = 5)
```

*Overall it looks like there can still be substantial fluctuation in log scores, but maybe slightly less than when scored on an absolute scale. Interestingly, it seems like the smoothing effect is stronger for deaths than for cases.*

```{r}
# create version of data where the true value is the true growth rate
# interesting question: does it make a difference, whether we transform it 
# to be the weekly growth rate, i.e. take the n-th root, where n is the horizon?

latest_truth <- fread(here("data", "weekly-truth-Europe.csv")) |>
  rename(last_known_true_value = true_value) |>
  mutate(forecast_date = target_end_date + 2) |>
  select(location, target_type, forecast_date, last_known_true_value) |>
  unique()

# unsure whether the pmax hack is the right way to do it
growth_data <- hub_data |>
  inner_join(latest_truth) |>
  mutate(true_value = true_value / pmax(last_known_true_value + 1),
         prediction = prediction / pmax(last_known_true_value + 1)) |>
  select(-last_known_true_value)
```

## Variance of scores vs. mean of scores across locations

```{r}
scores |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(target_type, location, scale) |>
  summarise(mean_wis = mean(interval_score), 
            sd_wis = sd(interval_score), 
            .groups = "drop_last") |>
  ggplot(aes(y = sd_wis, x = mean_wis)) + 
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", size = 0.5) + 
  geom_point() + 
  facet_wrap(~ target_type + scale, scale = "free") + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  # scale_x_continuous(labels = scale_fn) + 
  # scale_y_continuous(labels = scale_fn) + 
  scale_x_continuous(labels = scale_fn, trans = "log10") +
  scale_y_continuous(labels = scale_fn, trans = "log10") +
  labs(x = "Mean WIS", y = "Sd of WIS")

ggsave("output/figures/HUB-sd-vs-mean-scores.png", width = 7, height = 5)
```

```{r}
# negative binomial
# MASS::glm.nb(true_value ~ target_type, 
#        data = hub_data %>%
#          mutate(true_value = pmax(true_value, 0)), 
#        link = "identity")
```

## WIS vs. the totsl number of observed values
probably not needed anymore, as we have similar things

```{r}

# wis_vs_total(scores) + 
#   labs(y = "Mean WIS")
# ggsave("output/figures/HUB-mean-scores-vs-total.png", width = 7, height = 5)
# 
# wis_vs_total(scores, summary_fct = stats::sd) + 
#   labs(y = "Sd WIS")
# ggsave("output/figures/HUB-sd-scores-vs-total.png", width = 7, height = 5)

# wis_vs_total(scores, summary_fct = stats::sd) + 
#   labs(y = "Sd WIS") + 
#   scale_x_continuous(labels = scale_fn, trans = "log10") +
#   scale_y_continuous(labels = scale_fn, trans = "log10") 
# ggsave("output/figures/HUB-sd-scores-vs-total-log-log.png", width = 7, height = 5)

```

Test for a single example forecast to verify that rankings change
```{r}
# we see that rankings change even for a single forecast. So we need to check everything is still correct. 
# take one example
test_data <- hub_data |>
  filter(model %in% c("MUNI-ARIMA", "epiforecasts-EpiNow2"), 
         forecast_date == "2021-10-18", 
         target_type == "Cases", 
         location == "AT")

test_data |>
  score(metrics = c("interval_score")) |>
  summarise_scores()

test_data |>
  mutate(prediction = log(prediction), 
         true_value = log(true_value)) |>
  score(metrics = c("interval_score")) |>
  summarise_scores()
## indeed everything seems to be correct, but rankings still change FOR A SINGLE FORECAST. 
# It seems this is because the trade-off between calibration and sharpness changes? 
# need to check in what way
```

## Distribution of scores

```{r}
scores |>
  ggplot(aes(x = interval_score)) + 
  geom_density() + 
  facet_wrap(~ target_type + scale, scale = "free") + 
  scale_x_continuous(labels = scale_fn)
```

## Fit a model to data

Check fit of the log scale model
```{r echo=TRUE}
reg_log <- scores |>
  filter(scale == "log") |>
  glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 
      family = gaussian)
```

```{r}
# check distribution
performance::check_distribution(reg_log) |>
  plot()

# check normality of residuals
check_normality(reg_log) |>
  plot(type = "qq")

# heteroskedasticity of error terms - the plot takes extremly long to run
performance::check_heteroskedasticity(reg_log)
# performance::check_heteroskedasticity(reg_log) |>
#   plot()

# check influence of outliers - plot also takes very long to run
performance::check_outliers(reg_log)
# performance::check_outliers(reg_log) |>
#   plot()

```

---

## Comparison of the distribution of model rankings

```{r}
rankings <- scores |>
  select(-c(coverage_deviation, bias, ae_median, dispersion, underprediction, overprediction)) |>
  group_by(location, forecast_date, horizon, target_type, scale) |>
  mutate(ranking = rank(interval_score) / n()) |>
  ungroup()

rankings_wide <- rankings |>
  select(model, location, target_end_date, forecast_date, horizon,
         interval_score, scale, ranking, target_type) |>
  pivot_wider(names_from = scale, values_from = ranking, 
              id_cols = -interval_score)

plot_ranking_diffs <- function(
  rankings, 
  rankings_wide,
  target_type
) {
  type <- target_type
  # maybe replace this by the same ranking for all plots
  factor_levels <- rankings |>
    filter(horizon == 2, scale == "natural", target_type == type) |>
    group_by(model) |>
    summarise(ranking = mean(ranking)) |>
    arrange(-ranking) |>
    pull(model)

  # p_rankings <- rankings |>
  #   filter(horizon == 2) |>
  #   mutate(model = factor(model, levels = factor_levels)) |>
  #   ggplot(aes(y = model, x = ranking)) +
  #   ggdist::stat_halfeye() +
  #   facet_wrap(~ scale) +
  #   theme_scoringutils() +
  #   labs(y = "Model", x = "Normalised ranking")

  ranking_plot <- function(
    hor
  ) {
    p <- rankings_wide |>
      mutate(diff_ranking = natural - log) |>
      mutate(model = factor(model, levels = factor_levels)) |>
      filter(horizon == hor, 
             !is.na(model)) |>
      ggplot(aes(y = model, x = diff_ranking)) +
      geom_vline(linetype = "dashed", xintercept = 0, 
                 size = 0.6, color = "grey40") +
      ggdist::stat_pointinterval() +
      theme_scoringutils() +
      labs(y = "", x = "Differences in ranking")

    if (hor > 1) {
      p <- p +
        theme(axis.text.y = element_blank(), 
              axis.ticks.y = element_blank(), 
              axis.line.y = element_blank())
    }
    return(p)
  }

  ranking_plot(hor = 1) + ranking_plot(2) + ranking_plot(3) + ranking_plot(4) +
    plot_layout(nrow = 1)
}

# should probably have one plot for both cases and deaths since this is rankings
p_diff_cases <- plot_ranking_diffs(rankings, rankings_wide, 
                                   target_type = "Cases")

ggsave(filename = "output/figures/HUB-ranking-diffs.png", 
       plot = p_diff_cases,
       width = 9, height = 5)
```

# Growth Rate stuff - Looking at growth rates directly

## Scores of the growth rate across targets

This uses the Hub ensemble
```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  ggplot(aes(y = interval_score, x = target_type)) + 
  geom_violin(aes(fill = target_type), alpha = 0.2, color = NA) + 
  geom_boxplot(alpha = 0.5) + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ scale, scale = "free") + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn, trans = "log10") + 
  labs(y = "Interval score", x = "Target type")
```
*Scoring the growth rate makes things more comparable*

Mean scores: 
```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(scale, target_type) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") |>
  make_table()
```

## Scores for the growth rate across locations
```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble", 
         horizon == 2) |>
  group_by(target_type, location, scale) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") |>
  ggplot(aes(y = interval_score, x = reorder(location, -interval_score))) + 
  geom_bar(stat = "identity") + 
  scale_fill_brewer(palette = "Set1", name = "Forecast target") + 
  facet_wrap(~ target_type + scale, scale = "free") + 
  theme_scoringutils() + 
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scale_fn) + 
  labs(y = "Interval score", x = "Target type")
```

## Scores for growth rate over time
restrict to a few locations, e.g. GB, FR, ES (highest average scores there)

UK: 
```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble",
         location == "GB",
         horizon == 2) |>
  group_by(target_type, scale) |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale, scale = "free_y") + 
  theme_scoringutils() + 
  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
```

France: Looks like a data revision
```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble",
         location == "FR",
         horizon == 2) |>
  group_by(target_type, scale) |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale, scale = "free_y") + 
  theme_scoringutils() + 
  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
```

Spain: 
```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble",
         location == "ES",
         horizon == 2) |>
  group_by(target_type, scale) |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = forecast_date)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale, scale = "free_y") + 
  theme_scoringutils() + 
  labs(x = "Forecast date", y = "WIS relative to smallest observed value")
```

## Change of average scores for the growth rate for increasing forecast horizons
```{r}
scores_gr |> 
  filter(model == "EuroCOVIDhub-ensemble") |>
  group_by(scale, target_type, horizon) |>
  summarise(interval_score = mean(interval_score), 
            .groups = "drop_last") |>
  mutate(multiplicative_score = interval_score / min(interval_score)) |>
  ggplot(aes(y = multiplicative_score, x = horizon)) + 
  geom_line() + 
  geom_point() + 
  facet_grid(target_type ~ scale) + 
  theme_scoringutils() + 
  labs(x = "Forecast horizon in weeks", y = "Multplicative change in interval score")
```

## Distribution of scores growth rate

```{r}
scores_gr |>
  ggplot(aes(x = interval_score)) + 
  geom_density() + 
  facet_wrap(~ target_type + scale, scale = "free") + 
  scale_x_continuous(labels = scale_fn)
```

## Fit a model to growth rate data

Check fit of the log scale model
```{r echo=TRUE}
reg_gr <- scores_gr |>
  filter(scale == "natural") |>
  glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 
      family = gaussian)

reg_gr_log <- scores_gr |>
  filter(scale == "log") |>
  glm(formula = interval_score ~ 1 + model + location + target_type + horizon, 
      family = gaussian)
```

Check distribution
```{r}
# check distribution
performance::check_distribution(reg_gr_log) |>
  plot()

# check normality of residuals
check_normality(reg_gr_log) |>
  plot(type = "qq")

# heteroskedasticity of error terms - the plot takes extremly long to run
performance::check_heteroskedasticity(reg_gr_log)
# performance::check_heteroskedasticity(reg_log) |>
#   plot()

# check influence of outliers - plot also takes very long to run
performance::check_outliers(reg_gr_log)
# performance::check_outliers(reg_log) |>
#   plot()

```

Figure
Currently discontinuoued and just there to discuss it. 
that shows Variance vs. mean of the observations. Makes the point that the
data looks neg. binomial. 
```{r}
# plot with mean vs. variance --------------------------------------------------
plot_var_mean <- function(hub_data, log = FALSE) {
  df <- hub_data |>
    mutate("x" = true_value) |>
    dplyr::select(location, target_type, target_end_date, true_value, x) |>
    unique() |>
    group_by(location, target_type)

  f <- function(x) {x + x^2}

  if (log) {
    df <- mutate(df, true_value = log(pmax(true_value, 0) + 1))

    f <- function(x) {log(log(x) + log(x)^2 + 1)}
  }

  out <- df |>
  summarise(mean_value = mean(x), 
            sd_value = sd(true_value), 
            var_value = var(true_value)) |>
  ggplot(aes(y = var_value, x = mean_value)) +
  geom_point() + 
  facet_wrap(~ target_type, scales = "free") + 
  scale_x_continuous(trans = "log10", labels = label_fn) +
  scale_y_continuous(trans = "log10", labels = label_fn) +
  labs(y = "Var observations", x = "Mean observations") + 
  # geom_function(fun = f, linetype = "dashed", size = 0.5) +
  theme_scoringutils() + 
  geom_smooth(alpha = 0.2, size = 0.2)

  if (log) {
    out <- out +
      labs(y = "Var log(observations)", x = "Mean observations")
  }

  return(out)
}

p1 <- plot_var_mean(hub_data)
p2 <- plot_var_mean(hub_data, log = TRUE)

wis_vs_total <- function(scores, summary_fct = mean, fulldata = hub_data) {

  mean_values <- 
    fulldata |> 
    filter(!is.na(prediction)) |>
    group_by(location, target_type, target_end_date) |>
    summarise(true_value = unique(true_value)) |>
    group_by(location, target_type) |>
    summarise(mean_obs = mean(true_value))

  plot_df <- scores |> 
    filter(model == "EuroCOVIDhub-ensemble", 
           horizon == 2) |>
    group_by(location, target_type, scale, type_and_scale) |>
    summarise(wis = summary_fct(interval_score), 
              .groups = "drop_last") |> 
    inner_join(mean_values)

  plot_df |>
    group_by(target_type, scale) |>
    summarise(correlation = cor(wis, mean_obs)) |> 
    print()

  plot_df |>
    ggplot(aes(y = wis, x = mean_obs)) + 
    geom_point() + 
    facet_wrap(~ type_and_scale, scale = "free") + 
    theme_scoringutils() + 
    theme(legend.position = "none") + 
    scale_x_continuous(labels = scale_fn) +
    scale_y_continuous(labels = scale_fn) +
    labs(y = "REPLACE", x = "Mean observed value")
}

# log-log versions of that plot

scatter_wis_obs <- wis_vs_total(scores) + 
  labs(y = "Mean WIS") + 
  scale_x_continuous(labels = scale_fn, trans = "log10") +
  scale_y_continuous(labels = scale_fn, trans = "log10")

p1 / p2 / scatter_wis_obs + 
  plot_layout(heights = c(1, 1, 2))

ggsave("output/figures/HUB-mean-scores-vs-total-log-log.png", width = 7, height = 7)
```